### 2008-05-26 paper/paper.tex
Revised paper.

% TODO: ensure all references in paper.bib are cited in the paper!
% TODO: format checklist (see e-mail)
\documentclass{acm_proc_article-sp}

\usepackage{pslatex}
\usepackage{epsfig}
\usepackage{appendix}
\usepackage{float}
\usepackage{url}

\floatstyle{ruled}
\newfloat{program}{thp}{lop}
\floatname{program}{Program}

\begin{document}

\title{Evaluation of Cache-Oblivious Data Structures [draft-1]}

\numberofauthors{1}
\author{Maks Verver\\ \email{m.verver@student.utwente.nl}}

% Obligatory permission block
\toappear{
Permission to make digital or hard copies of all or part of this work
for personal or classroom use is granted without fee provided that
copies are not made or distributed for profit or commercial advantage
and that copies bear this notice and the full citation on the first
page. To copy otherwise, or republish, to post on servers or to
redistribute to lists, requires prior specific permission.

\textit{\small 9$^{th}$ Twente Student Conference on IT, Enschede, June, 2008}

Copyright 2008, University of Twente, Faculty of Electrical Engineering,
Mathematics and Computer Science}
% End of obligatory permission block

\maketitle

\begin{abstract}
In modern computer hardware architecture memory is organized in a hierarchy
consisting of several types of memory with different memory sizes, block
transfer sizes and access times. To reflect this situation more accurately
than traditional models, the cache-oblivious model has been proposed, and
several algorithms and data structures have been developed that perform
optimally in this model.
This paper presents the results of an empirical evaluation of several data
structures in a realistic scenario, and gives insight into the applicability
of cache-oblivious data structure in practice.
\end{abstract}

\keywords{cache efficiency, locality of reference, algorithms}

% TODO: more meaningful section titles?

\section{Introduction}
A fundamental part of theoretical computer science is the study of algorithms (formal descriptions of how computations may be performed) and data structures (descriptions of how information is organized and stored in computers). Traditionally, algorithms have been evaluated in a simplified model of computation. In this model it is assumed that a computer executes an algorithm in discrete steps. At each step it performs one elementary operation (e.g. comparing two numbers, adding one to another, storing a value in memory, et cetera). Each elementary operation is performed within a constant time. In this model, both storing and retrieving data values in memory is considered to be an elementary operation.

This model is close enough to the way computers work to be extremely useful in the development and analysis of data structures and algorithms that work well in practice. However, like every model, it is a simplification of reality. One of the simplifications is the assumption that storing and retrieving data in memory is a constant-time operation (i.e. the uniform memory model). Advancements in software and hardware design over the last two de\-cades have caused this assumption to be increasingly detached from reality.

One of the reasons this assumption fails is that modern operating systems employ virtual memory management, which allows (slower) disk-based storage to be dynamically substituted for (fast\-er) main memory when the latter is in short supply. Another reason is that as processor speeds have increased greatly, the time required to transfer data between processor and main memory has become a bottleneck for many types of computations. Hardware architects have added faster (but small) cache memory at various points in the computer architecture to reduce this problem, but this increases the differences in access time between different types of memory.

As a result, a modern computer system lacks a central memory storage with uniform performance characteristics. Instead, it employs a hierarchy of memory storage types. Figure \ref{fig-memhier} gives a typical example of such a hierarchy. The processor can directly manipulate data contained in its registers only. To access data in a lower level in the memory hierarchy, the data must be transferred upward through the memory hierarchy. Memory is typically transferred in blocks of data of a fixed size (although bulk transfers involving multiple blocks of data at once are also supported at the memory and disk level).

\begin{figure}
\centering
\includegraphics{memhier}
\caption{Schematic depiction of the memory hierarchy}\label{fig-memhier}
\end{figure}

In the memory hierarchy, every next level of storage is both significantly slower and significantly larger than the one above it and with increasing memory sizes, the block size increases as well. Table \ref{tab-memhier} gives an overview of typical memory sizes, block sizes, and access times. The values for the (general purpose) registers, L1 cache and L2 cache are those for the Pentium M processor as given in \cite{intel-opt}; other values are approximate.

\begin{table}
\begin{center}
\begin{tabular}{ l l l l }
\hline
\textbf{Type} & \textbf{Total Size} & \textbf{Block Size} & \textbf{Access Time} \\
\hline
Registers  &   32 bytes & 4 bytes & 1 cycle \\
L1 Cache   &   32 KB    & 64 bytes & 3 cyles \\
L2 Cache   &    1 MB    & 64 bytes & 9 cycles \\
RAM        &   ~2 GB    & 1 KB     & 50-100 cycles \\
Disk       & ~300 GB    & 4 KB     & 5,000-10,000 cycles \\
\hline
\end{tabular}
\caption{Sizes and access times of various types of storage}
\label{tab-memhier}
\end{center}
\end{table}

To conclude: the memory model used in real computer systems is quite a bit more complex than the uniform memory model assumed. Given this reality, the performance of many existing algorithms and data structures can be improved by taking the existence of a memory hierarchy into account. This has prompted research into new memory models that are more realistic. We will consider three classes of data structures and algorithms, depending on the assumptions that are made about the environment in their definition:
\begin{list}{}{}
\item \textbf{Cache-unaware} data structures and algorithms are designed with the traditional uniform memory model in mind, without considering the existance of a memory hierarchy, and may therefore not perform optimally in practice.
\item \textbf{Cache-aware} (or \textbf{cache-conscious}) data structures and algorithms are designed to perform optimally in the external memory model (described below) but require parametrization with some or all properties of the cache layout (such as block size or cache memory size).
\item \textbf{Cache-oblivious} data structures and algorithms are designed to perform optimally in the cache-oblivious model (explained below) which does not allow para\-metri\-zation with properties of the cache layout.
\end{list}

Plenty of cache-unaware and cache-aware data structures have been developed and these are also widely used in practice. Altough some research has been done on cache-oblivious data structures, they are seldom used in practice. The main contribution of this paper is to give some insight in the practical merit of cache-oblivious data structures.

\subsection{External Memory Model}
One of the earliest models to take the differences in memory access cost into account is the external memory model, described by Aggarwal and Vitter \cite{aggarwal1988ioc}. They make a distinction between internal memory (which is limited in size) and external memory (which is virtually unlimited). The external memory is subdivided into blocks of a fixed size and only entire blocks of data can be transferred between the two memories; additionally, consecutive blocks can be transferred at reduced cost (so-called bulk transfer). Although Aggarwal and Vitter focus on magnetic disk as a storage medium for external memory, the model can be generalized to apply to every pair of adjacent levels in the memory hierarchy. In that case, the possibility of bulk transfer may have to be dropped.

We will call algorithms that are designed to minimize the number of transfers in a two-level memory hierarchy ``cache-aware'' (as opposed to traditional ``cache-unaware'' algorithms) or ``cache-conscious''. These algorithms typically rely on knowledge of the block size to achieve optimal performance.

\subsection{Hierarchical Memory Model}
The external memory model has the limitation that it only describes two levels of storage, while we have seen that in practice the memory hierarchy contains more than just two levels. Even though the external memory model can be used to describe any pair of adjacent levels, a particular algorithm can only be tuned to one. Aggarwal, Alpern, Chandra and Snir \cite{aggarwal1987mhm} addressed this shortcoming by introducing a hierarchical memory model in which the cost of accessing a value at specific memory addresses is described by a non-decreasing function of these addresses, which means that accessing data at higher addresses can be slower than at lower addresses. This more closely resembles a real memory hierarchy, but assumes that the application has full control over which data is placed where and that data at all locations can be accessed directly, which is usually not the case in practice. As a result, the applicability of their model to the design and evaluation of \emph{practical} algorithms is limited.

\subsection{Cache-Oblivious Memory Model}
A different approach to generalizing the external memory model was taken by Prokop \cite{prokop1999coa} who proposed the cache-oblivious model, which consists of an infinitely large external memory and an internal memory of size M which operates as a cache for the external memory. Data is transferred between the two in aligned data blocks of size B. In contrast with the hierarchical memory model, the application does not have explicit control over the transfering of blocks between the two memories. Instead, it is assumed that an optimal cache manager exists which minimizes the number of block transfers over the execution the program. Additionally (and in contrast with the external memory model) the values of parameters like M and B are not known to the application, so they cannot be used explicitly when defining algorithms and datastructures. Of course, analysis in terms of memory transfers does involve these parameters, so the number of memory transfers performed is still a function of M and B (and other parameters relevant to the problem).

Algorithms that perform optimally in this model are called ``cache-oblivious'' and they distinguish themselves from cache-aware algorithms in that they cannot rely on knowledge of the block size or other specific properties of the cache configuration. The key advantage of this class of algorithms is that, even though they are defined in a two-level memory model, they are implicitly tuned to all levels in the memory hierarchy at once. It has been conjectured that these algorithms may therefore perform better than algorithms that are tuned to a specific level in the hierarchy only.

Demaine gives an introduction into the cache-oblivious memory model and an overview of a selection of cache-oblivious data structures and algorithms \cite{demaine2002coa}. He also motivates the simplifications made in the cache-oblivious memory model, such as the assumption of full cache associativity and an optimal replacement policy.

% Copy and shorten introduction from proposal
% -> Motivate research into cache-efficient algorithms
% -> Move "classification" subsection describing three models to front
% -> Describe each of the three models
% -> Introduce three data structures:
%     - note that hash tables have poor locality of reference
%     - note that btrees have good locality of reference but require pagesize
% -> Finally, present goal of the research (incorporate "Research Questions" in
% proposal)

% Marielle suggested:
%  - give the "standard" memory model an explicit name
%    (isn't "cache-unaware" enough?)
%  - move examples of cache oblivious data structures to front
%  - mention examples of cache unaware and cache aware algorithms
%    (at least Btree, tries and hash tables are mentioned)

% Note explicitly that cache-oblivious algorithms are relatively complex
% resulting in higher constant consts, so that the main question is wether
% the improved locality of reference compensates for the overhead incurred.

\section{Contributions}
Several cache-oblivious data structures and algorithms have been proposed. Complexity analysis shows that the proposed solutions are asymptotically optimal. However, in software engineering practice we are not only interested in computational complexity, but also in the practical performance of data structures and algorithms. Indeed, many algorithms that have suboptimal computational complexity are actually widely used because they perform well in practice (for example, sorting algorithms like Quicksort and Shell sort) and the converse is true as well: some algorithms, although theoretically sound, are unsuitable for practical use because of greater memory requirements, longer execution time, or difficulty of implementation (for example: linear-time construction of suffix arrays is possible, but in practice slower alternatives are often preferred that are easier to implement and require less memory).

This raises the question whether cache-oblivious data structures are actually preferable to traditional data structures in practice. To determine if cache-oblivious data structures have practical merit, empirical performance data is required, which is scarce as existing research has focused mainly on theoretical analysis. This paper addresses the question by reporting on the performance of newly implemented (but previously described) cache-oblivious data structures and their traditional counterparts (both cache-aware and cache-unaware data structures).

\section{Related Work}
Many data structures and algorithms have been analyzed in the cache-oblivious model. Several new data structures and algorithms have been developed that perform optimally in this mo\-del as well. Prokop presents asymptotically optimal cache-oblivious algorithms for matrix transposition, fast Fourier transformation and sorting \cite{prokop1999coa}.

Bender, Demaine and Farach-Colton designed a cache-oblivious data structure that supports the same operations as a B-tree \cite{bender2005cob} achieving optimal complexity bounds on search and nearly-optimal bounds on insertion. This data structure is mostly of theoretic interest, as it is fairly complex to implement and incurs a relatively large overhead in practice.

Later, Bender, Duan, Iacono and Wu simplified the data structure \cite{bender2004lpc} while preserving the supported operations and complexity bounds and adding support for finger searches. The latter data structure consists of a sparse array (stored in a linear amount of memory) with a static tree in van Emde Boas layout (named for similarity to a different data structure proposed by van Emde Boas et al. \cite{vanemdeboas1976dai}).

Several data structures were proposed by Rahman, Cole and Raman \cite{rahman2001opd} amongst them a cache-oblivious exponential search tree with a similar structure as the static search tree proposed by Prokop. In their experimental results the cache-oblivious tree performs worse than the (non-oblivious) alternatives. Nevertheless, they conclude that \textit{``cache-obli\-vious data structures may have significant practical importance''}.

Askitis and Zobel \cite{askitis2005ccc} propose a way to optimize seperate-chaining hash tables for cache-efficiency by storing the linked lists that contain the contents for a bucket in contiguous memory. Their experiments show a performance gain over traditional methods, especially when the hash table is heavily loaded.

Vitter presents a theoretical survey of algorithms evaluated in a parallel disk model \cite{vitter2001ema}, which is a refinement of the external memory model described by Aggarwal and Vitter, but only allows parallel transfer of multiple blocks from different disks, which is more realistic. Unfortunately, his survey lacks empirical results.

Olsen and Skov evaluated two cache-oblivious priority queue data structures in practice \cite{olsen2002coa} and designed an optimal cache-obli\-vious priority deque. Their main result is that although the cache-obli\-vious data structures they examined make more efficient use of the cache, they do not perform better than traditional priority queue implementations.

\section{Methods}
In order to gather empirical data, the research approach must be made more concrete. We will need to limit ourselves to a specific class of data structures, since algorithms and data structures offering different functionality cannot be compared in a meaningful way. Furthermore, we will need to select a proper scenario in which the data structures are evaluated, as conclusions on the practical merit of the data structures depend on the degree to which the test scenario is realistic.

Furthermore, we will need to define what practical performance is. We will measure two properties: primarily, the total running time of our test application, and secondarily the memory used. The rationale for selecting these metrics is that if the data structures perform identical functions and plenty of memory is available, the only observable difference in running a program using different data structures will be the run time. Memory use is of secondary interest because in practice memory may be limited, which would preclude the use of certain data structures that require a lot of memory to function.

\subsection{State Space Search}
The test application that we use to gather performance data is an implementation
of a state space search algorithm.
This is a suitable scenario for two reasons.
First, it is commonly used as a practical
component of formal methods for software verification, and therefore good
algorithms are of great practical significance. Second, as we will explain
below, the performance of state space search algorithms depends for a large
part on the performance of the data structures that are used to implement
them; therefore, research into efficient data structures is of particular
interest to this application.

State space search can be used to verify the correctness of software programs. For this purpose, programs are first modeled using a formal language, that is also used to specify properties of the program that should hold during its
execution. An executing program can be in a (possibly infinite) amount of states,
one of which is usually designated the \emph{initial state}. If a transition
from one state to another is possible (according to the rules of the formal
language used) the latter state is said to be a \emph{successor state} of the
former.
Generating the successors of a particular state is also called
\emph{expanding} the state. Of course, the the execution model
must include some form of non-determinism to allow more than one successor to
exist for a single state.
In practice, this non-determinism usually comes from processes that execute in
parallel, where the precise interleaving of the execution of
instructions in these processes is non-deterministic, unless synchronization
primitives (such as channels, semaphores, atomic execution blocks, et cetera)
are used to enforce a particular ordering.

The set of all states reachable by transitions from the initial state is called
the \emph{state space} of a model, and it is the goal of a state space search
algorithm to generate all of these states, in order to check that desired
properties hold in all of them. For this approach it is often required that
the state space be finite (for obvious reasons).

\begin{program}
\begin{verbatim}
Queue queue = new Queue;
Set generated = new Set;
queue.insert(initial_state);
while (!queue.empty()) {
    State state = queue.extract();
    for (State s : successors(state)) {
        if (generated.contains(s) == false) {
            generated.insert(s);
            queue.insert(s);
        }
    }
}
\end{verbatim}
\caption{Pseudo-code for a simple state search algorithm.}
\label{prog-search}
\end{program}

The outline of a state space search algorithm is given in Program
\ref{prog-search}. Note that in addition to the initial state and a function
to generate successors, a queue and a set data structure are used. The queue
holds states that have been generated but not yet expanded and is used to
determine what state to expand next. The set holds all states that have been
generated so far and is used to prevent a single state from being expanded
more than once.

Depending on how states are inserted and extracted from the queue, the search
may be breadth-first (meaning that all states that are reachable in $N$ steps
from the initial state are expanded before any states that require more than
$N$ steps) or depth-first (meaning that states that are generated most recently
by the successor function are expanded first). We will use breadth-first
search, which means the queue operates by a first-in first-out principle.

From the pseudo-code it is clear that the state space search algorithm does not
accomplish much by itself; instead, the real work is done by the successor
function and the queue and the set datastructures. Queues can easily be
implemented efficiently (adding and removing states takes O(1) time).
The efficiency of the successor function depends on the execution model used,
but is typically linear in the number of states produced. In practice,
therefore, the bottleneck of the algorithm is the set data structure.

\subsection{Experimental Framework}
For our experimental framework, we need a collection of formal models that are
representative of those typically used for formal verification, and a way to
execute them. For the first part, we have looked at Spin \cite{spin}, a widely
used model checking tool. Spin uses a custom modeling language (called
PROMELA) to specify models and is distributed with a collection of example
models that are suitable for our experiments.

For the execution of these models we will use the NIPS VM \cite{weber2007evm}, a
high-performance virtual machine for state space generation that is easily
embeddable in our framework. Although the NIPS VM only executes bytecode in a
custom format, a PROMELA compiler is also available to generate the required
bytecode from the Spin models.\cite{nipsvm} The NIPS VM is preferred to the
Spin tools because it was designed to be embedded in other programs and as
such is more easily integrated in our framework.

The NIPS VM represents program state as a binary string; the size of the state
depends amongst other on the number of active processes in the program (which may change over the execution of the program). A typical state size is in the order of a few hundred bytes.

\subsection{Data Structures}
In the introduction we identified three classes of data structures.
For our evaluation we have implemented a single representative data structure
for each class:
\begin{itemize}
\item \textbf{Cache-unaware}: hash tables. Hash tables are widely used in practice and noted for good performance if the data set fits in main memory, although they have also been used as an index structure for data stored on disk.
\item \textbf{Cache-aware}: B-trees. The B-tree is the standard choice for storing ordered data on disk, and depends on a page size being selected that corresponds with the size of data blocks that can be efficiently transferred.
\item \textbf{Cache-oblivious}: the data structure proposed by Bender, Duan, Iacono and Wu. Since it provides functionality comparable to that of a B-tree, this seems like a fair candidate for a comparison. For brevity, this data structure will be referred to as a \emph{Bender set}.
\end{itemize}

Cache-oblivious data structures are not yet commonly used in practice and to
our knowledge there are no high-quality implementations publicly available.
The Bender set will therefore have to be implemented from scratch.

In contrast, both hash tables and B-trees are widely used and there are several
high quality implementations available as software libraries. It is, however,
undesirable to use existing libraries for a fair comparison, for two reasons. First, many existing implementations support additional operations (such as removal of elements, locking and synchronization, atomic transactions, et cetera) which are not used in our experiments, but which may harm the performance of these implementations. Second, many established libraries have been thoroughly optimized while our newly implemented data structures have not. This may give the existing libraries an unfair advantage.

In an attempt to reduce the bias caused by differences in functionality and
quality between existing and newly developed libraries, all data structures
used in the performance evaluation have been implemented from scratch.

\subsubsection{Set Operations}
Dynamic set data structures may support various operations, such as adding and removing elements, testing if a value is an element of the set, finding elements nearby, succeeding or preceding a given value, counting elements in a range, et cetera. However, our state space search algorithm only needs two operations: insertion of new elements (in a set that is initially empty) and testing for the existence of elements. In fact, these two operations can be combined into a single operation. We call this operation \verb#insert(S,x)#. If an element $x$ does not exist in $S$, then \verb#insert(S,x)# inserts it and returns $0$. If $x$ is already an element of $S$, \verb#insert(S,x)# returns $1$ and no modifications are made. The inner loop of our state space algorithm can then be rewritten as follows:
\begin{verbatim}
    for (State s : successors(state)) {
        if (generated.insert(s) == 0) {
            queue.insert(s);
        }
    }
\end{verbatim}

We now have a single operation that must be implemented by all set data structures. Recall that the values to be stored are virtual machine state descriptions, which are variable-length binary strings. All data structures must therefore support storing strings of varying length efficiently.

\subsubsection{Common Implementation Details}
For memory management, the data structures make use of memory mapped files.
The B-tree and hash table implementations use a single file for all data storage;
the Bender tree uses several files (as discussed below) and some main memory.

The use of memory mapped files means that the operating system is in charge of
when and which pages are transferred between main memory and disk storage; the
application assumes no control over this policy. This is in accordance with the
cache-oblivious model, that assumes no control over the cache is possible.
Additionally, memory mapped files bypass the C library's allocation functions
and give the implementor complete control over how data is ordered in (virtual)
memory.

\subsubsection{Hash Table Implementation}
In its simplest form, a hash table consists of an
index array (the \emph{index}) with a fixed number of slots. A \emph{hash
function} is used to map key values onto slot numbers. If the slot for a value
is empty, it may be inserted there. Queries for the existance of an element in
the hash table similarly see if the queried value is stored at its designated
slot.

When several values are inserted, some values may map to the same slot,
which is problematic if each slot can only store one value. There
are a lot of different ways to resolve this collision problem; we use
\emph{separate chaining} which means that slots do not store values directly,
but instead each slot stores a pointer to a singly-linked list of values that
map to that slot. This particular implementation technique is well-suited to
the scenario where values may have different sizes (as the slots only need to
store a fixed-size pointer and not a variable-size value) and maintains
relatively good performance when the number of values stored exceeds the size
of the index array \cite{sedgewick1998ac}.

% TODO: picture?

Our hash table implementation uses a fixed size index which must be specified
when creating the hash table, as this simplifies the implementation
considerably. The index is stored at the front of the file, after which values
are simply appended in the order in which they are added. Note that we do not
support removing elements from the hash table, which means we do not have to
deal with holes that would otherwise occur in the stored file. For our
experiments the FNV-1a hash function \cite{noll2004fnv} is used (modulo the
size of the index) to map values to slots.

\subsubsection{B-tree Implementation}
The B-tree data structure was first proposed by Bayer and McCreight
\cite{bayer1970oam} and is widely implemented and often described in
textbooks. Our implementation is based on the description by Kifer, Bernstein
and Lewis \cite{kifer2006dsa}.

B-trees are similar to other search tree data structures in the sense that
they store ordered values hierarchically in such a way that each subtree store
values that are close together.
A key property of B-trees is that they do not fix the number of children per node, but instead organize data in pages of a fixed size, each containing
as many values as will fit. This makes them especially suitable for purposes
of on-disk storage where reading and writing data in a few large blocks is
relatively cheap compared to accessing several smaller blocks.

B-tree pages are ordered in a tree structure. Values in a page are stored
in order, and pointers are added that originate between a pair of adjacent
values, and point to a subtree containing only values that fall in between those
values. The resulting tree is typically very shallow, which is benefical,
as the number of pages that need to be retrieved (worst case) is equal to
the height of the tree.

When a new value does not fit in the page where it is to be inserted,
the page is split into two pages; the median value from the old page is
moved to the parent page, and two new pages (containing the values less
than respectively greater than the median) are created. When the top-most
page needs to be split, a new (empty) top-level page is created and the height
of the B-tree is increased by one. As a result, all leaf nodes in a B-tree are
at the same depth, and all pages stored are at least half full.

% TODO: picture and description

B-trees can easily support the insertion of variable-length values, as long
as individual values fit in the size of a single page. However, there is no
obvious way to store values that are larger than the size of a single page.
Production-quality implementations handle this case separately. Our our
implementation does not support this requires stored values to
be smaller than the pagesize. Additionally, deletion of items has not been
implemented.

\subsubsection{Bender Set Implementation}
The first implementation challenge for the Bender set is that the description
given by Bender et al. assumes that all values stored in the set are of a
fixed size, which is not the case in our experimental framework. To work
around this, we create several separate instances of the Bender set with
different value sizes which are powers of two. When a value is to be inserted,
its size is rounded up to a power of two and it is inserted in the
corresponding set instance. This ensures that the amount of spaces wasted
remains below 50\% while still allowing values of various sizes to be inserted.
In the following description, we will assume that values are of a fixed size.

The implementation of the Bender set consists of two parts: a sparse array
storing the values in order, and a binary tree stored in van Emde Boas
layout (similar to the one proposed by Prokop) that is used to find values
efficiently.
Both the tree and the array may contain special \emph{empty values}.
A depiction of a Bender set storing four values (with a capacity for eight) is
given in Figure \ref{fig-bender-set}.

\begin{figure}
\centering
\includegraphics{bender-set}
\caption{A depiction of a Bender set}\label{fig-bender-set}
\end{figure}

A Bender set is created with a capacity $C$ that is a power of two.
Initially, the array stores only empty values and is partitioned into
\emph{windows} on several \emph{levels}.
On the highest level, there is a single window of size $C$. Each subsequent level
has twice as many windows of half that size, and the lowest level has size
$log_2 C$ (rounded up to the next power of 2).
For each level a maximum \emoph{density} is chosen, with the lowest level having
density 1, the highest level having density $\delta$, and the density for
the intermediary levels obtained by interpolating linearly between 1 and
$\delta$.

The top-level density must be a number between $0$ and $1$; the optimal value
depends on practical circumstances such as available memory and the relative
cost of rebuilding the data structures.
Whenever the fraction of non-empty values stored in a window (the window's
\emph{population count} divided by the size of the window) exceeds its maximum
density, the window is said to be \emph{overflowing}.
Overflows are resolved when a higher-level window is \emph{rebalanced},
meaning that the values in the window are redistributed evenly over the window,
which will cause some of the values to be moved out of the lower-level windows
which are too full.

To insert a value, the binary tree is used to find a position in the array
where the value may be inserted (just before its successor). Then, the windows
overlapping this position are considered from bottom to top, and the first
window that can support another element without overflowing is selected, and
then rebalanced (thereby resolving the overflow in all lower level windows).

In the case that all windows overlapping the target position would overflow upon
insertion of the new element, the capacity of the data structure is doubled,
a new top level is created and then the entire array is rebalanced. This is how
the set grows when the number of elements stored comes too close to the storage
capacity. Since changing the capacity causes all values in the array to be moved
and requires that the entire index tree is recreated, this is an expensive
operation; however, it occurs only infrequently.

According to this description (which follows the paper by Bender et al) a window
is rebalanced every time an element is inserted. As an optimization, our
implementation does not always rebalance a window. When there is free space
between the successor and predecessor of the value to be inserted, the value
is simply inserted in the middle. In our experiments, this optimization
yielded an reduction in execution time.

One thing that is not specified in the paper by Bender et al, is how the
population count for windows is kept. The simplest option is to keep no such
information and simply scan part of the array whenever a population count is
required. Another extreme is to keep population counts for all windows on all
levels, and update these counts whenever values are inserted or moved. In
our experiments a compromise seemed to work best: keep population counts only
for the lowest-level windows, and recompute the counts for higher-level windows
when required. This prevents a lot of recomputation, while keeping the additional
costs of updating low.

Finally, the Bender set uses a complete binary tree as an index data structure
to efficiently find the successor element for a given value. The tree is
stored in memory in van Emde Boas lay-out (as described by Prokop) to allow
efficient cache-oblivious traversal.

In the tree, the leaf nodes (when numbered from left to right) store the same values
as the sparse array (including empty values).
Interior nodes store the maximum value of their two children (or an empty value,
if both children store an empty value).
This tree can be used in a similar way as a binary search tree: by examining the
two children of a node, we can determine wether a searched-for value lies in the
left or right subtree.

Note that the \emph{structure} of the tree is static and only changes when the
capacity of the set is increased (in this case, the tree is recreated from scratch).
Of course, the values stored in the tree have to be updated when the corresponding
elements of the array change.

\subsection{Measurements}
Measuring the memory usage of a process can be done by retrieving how much of the
address space has been allocated by a process; called the \emph{virtual set
size}. This does not necessarily correspond one-to-one with memory being allocated for the process, but since the majority of the memory used by the
data structures is mapped at exactly one location, this metric is suitable
for our experiments.

There are several ways of measuring the time a process takes to execute.
The simplest is counting the number of seconds elapsed since the start of the
program (called \emph{wall clock time}).
This has the disadvantage that concurrently executing processes affect the runtime
of the process being measured, so on a busy system the reported runtime may vary
over different executions.

A different way to measure time is using the statistics the kernel collects
of how many seconds the processor spends executing instructions for the process
(both in user space and in kernel space). These values are independent of
what other processes are doing, which makes them more consistent across
multiple executions. However, these values do not account for the time the
system spends waiting (for example, for data to read from disk)
or the time spent by other processes on behalf of the executing process
(for example, the kernel swap daemon on Linux). Since these factors may have
a large effect on the total running time of the algorithm, it is undesirable
to leave these out.

Therefore, we decided to measure wall clock time. To deal with variations
across multiple executions, each experiment is run three times, and the median
value is used for further analysis.

\subsubsection{Framework Overhead}
Finally, there is another factor that must be taken into account: the test
framework uses several components (such as the NIPS VM and a queue) that are
not being evaluated, yet which do contribute to the memory use and runtime
of the process. In order to discount these factors, tests are performed using
a mock set implementation that gives a near-minimal overhead. This works
by first running with a real set implementation and logging all results
(the return values of the \verb#insert(S,x)# calls) to a file, and then
running a second time while reading the stored values. In the second run,
the mock set implementation does not have to actually store or retrieve
data and consequently has negligible memory and runtime overhead.

In the results below, all values are reported relative to the values obtained
using the mock set implementation, which means the overhead of the test
framework is removed from the results.

\subsubsection{System Configuration}
The experiments where performed on a 64-bit Linux (kernel 2.6) system
with an Intel Xeon E5320 processor (1.86 GHz, 4MB cache) and 8GB of main
memory and 2GB swap file. Although the system is a multiprocessor system
(and the E5320 is a dual-core processor) all code is single-threaded so
only a single core is used when executing tests.

% Describe data set used; per model:
%   average state size?
%   total size?
%   ratio of insertions/queries?

The following models are used for benchmarking:
\begin{itemize}
\item\textbf{Eratosthenes} is a parallel implementation of the sieve
of Eratosthenes which is used to find prime numbers upto a given maximum.
It has a single parameter $N$: only integers from 2 to $N$ (exclusive) are
tested. A new process is created for every prime number found, and as a result
the state space can only by exhaustively searched for relatively small values
of $N$ (e.g. $N < 40$). Because processes are dynamically created, state size
increases during execution.

\item\textbf{Leader2} is a leader election algorithm with a configurable number
of processes competing for leadership ($N$). Since a constant number of
processes will be created and these processes cannot make progress until all
of them are created, almost all states have the same size.

\item\textbf{Peterson\_N} is Peterson's solution to the multi-process mutual
exclusion problem (using global variables instead of channels for
communication) with a configurable number of processes ($N$). The processes
are created atomically, so the state size is constant after initialization.
\end{itemize}

\begin{table}
\begin{center}
\begin{tabular}{ l l l l }
\hline
\textbf{Model} & \textbf{Parameters} & \textbf{Iterations} & \textbf{Transitions} \\
\hline
Eratosthenes & $N=40$ &  1,019,960 &  4,923,218 \\
Leader2      & $N=5$  &  5,950,945 & 23,856,363 \\
Peterson\_N  & $N=4$  & 10,000,000 & 37,434,411 \\
\hline
\end{tabular}
\caption{Properties of test cases used}
\label{tab-cases}
\end{center}
\end{table}

In Table \ref{tab-cases} an overview is given of the parameters of the models used
and the resulting properties of the state space search.
Note that the number of transitions relative to the number of iterations indicates
the percentage of insertion operations that results in a new element being added
to the set, which ranges approximately from 4 to 2.7 non-updating queries per
insertion.

% Possible additional cases:
% - mobile1
% - pftp
% - snoopy
% - sort

\section{Results}

TODO: describe parameters of data structures
Explain results for B-tree (higher page size packs more data), hash table (larger index, less collissions, cache miss anyway)

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-queuesize}
\caption{Queue size}
\label{fig-quesize}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-transitions}
\caption{Transitions}
\label{fig-transitions}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-eratosthenes-bender_wctime}
\caption{Bender set execution time on Eratosthenes}
\label{fig-eratosthenes-bender_wctime}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-eratosthenes-bender_vss}
\caption{Bender set memory usage on Eratosthenes}
\label{fig-eratosthenes-bender_vss}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-eratosthenes-hash_wctime}
\caption{Hash table execution time on Eratosthenes}
\label{fig-eratosthenes-hash_wctime}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-eratosthenes-hash_vss}
\caption{Hash table memory usage on Eratosthenes}
\label{fig-eratosthenes-hash_vss}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-eratosthenes-btree_wctime}
\caption{B-tree execution time on Eratosthenes}
\label{fig-eratosthenes-btree_wctime}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-eratosthenes-btree_vss}
\caption{B-tree memory usage on Eratosthenes}
\label{fig-eratosthenes-btree_vss}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-leader2_wctime}
\caption{Execution times on Leader2}
\label{fig-leader2_wctime}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-leader2_vss}
\caption{Memory usage on Leader2}
\label{leader2_vss}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-peterson_wctime}
\caption{Execution time on Peterson\_N}
\label{fig-peterson_wctime}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=83mm]{fig-peterson_vss}
\caption{Memory usage on Peterson\_N}
\label{fig-peterson_vss}
\end{figure}

% Present results!
% Tables, graphs, etc.
% What to include:
%   queue size over iterations
%       for each program
%
%   insertions per query:
%       for each program
%
%   per program:
%       wctime over iterations
%          for each set
%       memory over iterations
%          for each set

\section{Discussion}
% Explain results, draw conclusions.
% - relate to insert/test ratio
% - relate to hardware (not many levels of hierarchy)

\section{Conclusion}
The experimental results clearly show that in terms of both memory usage and
execution time, the cache-oblivious data structure proposed by Bender, Duan,
Iacono and Wu is outperformed by traditional data structures in our test scenario.
The advantage of more cache-friendly behaviour does not appear to be large enough
to compensate for the increased complexity of the data structure, which results in
high ermemory requirements and increased computational overhead.
If the data structure does have asymptotic performance benefits, then realistic
work loads on current hardware systems are not enough to reveal this.
These findings are consistent with earlier results, such as obtained by
Rahman, Cole and Raman, and Olsen and Skov (see Related Work above).

Of course, this does not prove definitely that cache-oblivious data structures
are entirely without merit. It should be noted that the experiments performed
are limited in scope: only a single cache-oblivious data structure has been
examined, in a single test scenario, on a single hardware
and software platform. It is possible that in a different setting,
different cache-oblivious data structures do compare favorably to their traditional
counterparts. However, since the observed differences in performance are fairly
large, it is unlikely that small changes are enough to close the performance gap.

We hope that future research will focus not only on the theoretical performance
of newly developed algorithms and data structures, but also on the performance
in practice. Theoretical results are invaluable, but to be of greatest
significance newly developed data structures and algorithms should have
demonstrable practical merit as well.

\bibliographystyle{plain}
\bibliography{paper}
\end{document}